### 2024-06-16 basic-statistics.qmd
20240616 - statistical and practical significance

where $T$ is the test statistic of interest (e.g., $t$-, $r-$, $F$-, or $B$-coefficient, depending on the test), and $t$ is the observed test statistic.

@@

@@

>>>>

where $T$ is the test statistic of interest (e.g., $t$-, $r-$, $F$-, or $B$-coefficient, depending on the test), $t$ is the observed test statistic, and $Pr$ is the probability of the event.

@@

In science, we typically consider an effect to be *statistically significant* if the *p*-value is less than .05.
That is, there is a small chance (5%; or 1 in 20 times) that we would observe an effect as extreme as (or more extreme than) the effect observed, if the null hypothesis were true.
Statistical significance involves the *consistency* of an effect/association/difference; it suggests that the association/difference is reliably non-zero.
However, just because something is statistically significant does not mean that it is important.
In addition to statistical significance, it is also important to consider practical significance.

@@

*Practical significance* deals with how important the effect/association/difference is.
It is based on the magnitude of the effect, called the *effect size*.
Effect size can be quantified in various ways including:

- Cohen's $d$
- Standardized regression coefficient ($\beta$)
- Correlation coefficient ($r$)
- Coefficient of determination ($R^2$)
- Eta squared ($\eta^2$)
- Partial eta squared ($\eta_p^2$)

| Effect Size Index                             | Small        | Medium       | Large        |
|-----------------------------------------------|--------------|--------------|--------------|
| Cohen's d                                     | $\ge |.20|$ | $\ge |.50|$ | $\ge |.80|$ |
| Standardized regression coefficient ($\beta$) | $\ge |.10|$ | $\ge |.24|$ | $\ge |.37|$ |
| Correlation coefficient ($r$)                 | $\ge |.10|$ | $\ge |.24|$ | $\ge |.37|$ |
| Coefficient of determination ($R^2$)          | $\ge .01$   | $\ge .06$   | $\ge .14$   |
| Eta squared ($\eta^2$)                        | $\ge .01$   | $\ge .06$   | $\ge .14$   |
| Partial eta squared ($\eta_p^2$)              | $\ge .01$   | $\ge .06$   | $\ge .14$   |

: Effect Size Thresholds for Small, Medium, and Large Effect Sizes. {#tbl-effectSize Thresholds}
